---
title: "Homework assignment 05"
output: github_document
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r}

library(gapminder)
library(forcats)
library(ggplot2)
library(tidyr)
library(dplyr)
library(plotly)
library(readr)


```
#Part 1: Factor Rearrangement

##Drop Oceania.
Some information about the data set prior to change:
```{r}

# number of rows and levels of data set
(nrow(gapminder))
(nlevels(gapminder))
```
Dropping Oceania:

```{r}
gap_drop <- gapminder %>% 
  filter(continent != 'Oceania')

#display same information after dropped factor
(nrow(gap_drop))
(nlevels(gap_drop$continent))
```

Unused levels are still present. We can remove these by using $\texttt{droplevels()}:

```{r}
gap_drop_levels <- gap_drop %>% 
  droplevels()

#display same information after dropped unused levels
(nrow(gap_drop_levels))
(nlevels(gap_drop_levels$continent))
```

##Reorder the levels of country or continent

The data is by default ordered alphabetically. Let us instead order the countries by minimum population using $\texttt{fct_reorder}:

```{r}
reordered_cntry <- fct_reorder(gap_drop_levels$country, gap_drop_levels$pop, min) 

reordered_cntry %>% 
  levels() %>% 
  head(5) %>% 
  knitr::kable()
```
We can check whether this is correct by manually calculating the minimum population through the years:
```{r}
gap_drop_levels %>% 
  group_by(country) %>% 
  summarize(min_p = min(pop)) %>% 
  arrange(min_p) %>% 
  head(5) %>% 
  knitr::kable()

```
Which yields the same answer.


Let us order $\texttt{continent}$ by life expectancy:

```{r}

reordered_cont <- fct_reorder(gap_drop_levels$continent, gap_drop_levels$lifeExp, min) 

reordered_cont %>% 
  levels() %>% 
  head(5) %>% 
  knitr::kable()
  
```

And manually:

```{r}
gap_drop_levels %>% 
  group_by(continent) %>% 
  summarize(min_e = min(lifeExp)) %>% 
  arrange(min_e) %>% 
  head(5) %>% 
  knitr::kable()
```


#Part 2: File I/O

Let us first create a variable we can experiment with:

```{r}
exp_data <- gapminder %>% 
  mutate(continent_redord = fct_reorder(continent, lifeExp, min)) %>% 
  group_by(continent_redord) %>%
  summarize(min_lifeExp = min(lifeExp))

str(exp_data)
exp_data
  
```


##Write_csv()/read_csv()

Let us first try to write the data to file

```{r}
write_csv(exp_data, "test_write_to_csv.csv")

```

And then to read from file:

```{r}
data_csv <- read_csv("test_write_to_csv.csv")
data_csv
```
  We observe that the new column $\texttt{continent_redord}$ no longer is a factor, but a character. 
  
##saveRDS()/readRDS()
  
  Let us try to do the same thing, using the saveRDS() and readRDS():
  
```{r}
saveRDS(exp_data, "test_save_to_rds.rds")

```

And then read from .rbs:

```{r}
data_rds <- readRDS("test_save_to_rds.rds")
data_rds

```

##dput()/dget()

Can also use dput()/dget():

```{r}
dput(exp_data, "data_txt.txt")

```
```{r}
data_txt <- dget("data_txt.txt")
data_txt
```

#Part 3: Visualization design


Let us make a figure based on some basic functions. Want to make a plot that shows how life expectancy has changed throughout the years per continent:

```{r}
lexp_cont <- gapminder %>% 
  ggplot(aes(year, lifeExp)) +
  geom_point() +  #just to make things harder to interpret. Might as well use geom_smooth or geom_line
  aes(color = continent) + 
  facet_wrap(~continent) #showing all continents
lexp_cont
```


Instead, let us use some new techniques to obtain a nicer and more informative plot.

Let us first filter out some of the information that might be useful. We want to display some interesting facts about the life expectancy, like mean, max and min for each of the continents.
```{r}

filtered_data <- gapminder %>% 
  group_by(year, continent) %>% 
  summarize(min_lexp = min(lifeExp),
            max_lexp = max(lifeExp),
            mean_lexp = mean(lifeExp)) %>% 
  gather(key = "lexp_att", value = "lexp_val", min_lexp, max_lexp, mean_lexp)

lexp_new <- filtered_data %>% 
  ggplot(aes(year, lexp_val, group = lexp_att, color = lexp_att)) +
  scale_y_log10() + 
  geom_point() +
  geom_line() +
  facet_wrap(~continent) + 
  scale_color_discrete("Life expectancy distributions") +
  ylab("Life Expectancy")

lexp_new


```

We see that the way the data is displayed is more informative this way, with lines showing the change in life expectancy per year per continent. This way, the trends are more visible and interpretable. 

##Plotly


Differences:

-   $\texttt{Plotly}$ creates reactive plots where you would be able to hover your mouse over the plots to see data values and information, zoom in and out of regions as well as better customization of plots, e.g. color arrangements.
- ggplots may also be converted into plotly plots. 
- There is also a floating toolbar for additional options and interactions.

This means that $\texttt{plotly}$ in general is ment for making web-based graphs.   

Let us see how we can apply the plotly function to our new plot:

```{r}

#not viewable in document
ggplotly(lexp_new)
```

#Part 4: Writing figures to file

##Use of ggsave()

To save figures to a file, we could use the function $\texttt{ggsave()}$. This function has several parameters, like $\texttt{width}$, $\texttt{height}$, $\texttt{scale}$, $\texttt{dpi}$, making it possible to tune the image one wants to save to file. When desired plot is not specified, last plot displayed will be saved. 

##Saving an arbitrary plot

```{r}


#Save old plot

ggsave("./ggsave_figs/old_plot_img.png", plot = lexp_cont)

```


##Dimensions

```{r}


#default would save the file with the same dimensions as the original plot

ggsave("./ggsave_figs/default_img.png")


#fixed dimensions height and width:
ggsave("./ggsave_figs/fixed_dim_img.png", width = 10, height = 15)

#scale

ggsave("./ggsave_figs/scale_img.png", scale = 1.5)


```

##Resolution

```{r}
#fixed resolution, high:

ggsave("./ggsave_figs/high_dpi_img.png", dpi = 400)

#fixed resolution, low:

ggsave("./ggsave_figs/low_dpi_img.png", dpi = 10)


```

##Graphics devices

```{r}

#Vector format
ggsave("./ggsave_figs/eps_img.png", device = "eps")

#Bmp format
ggsave("./ggsave_figs/bmp_img.png", device = "bmp")



```

